The journey of Multiple Sequence Alignment

### 2021-08-02 MSA_Notum_journey.Rmd
Hs PORCN porcupine

## import fasta file with all sequences
import the sequences and filter them for the 10 best hits from the previous results in order to make the multiple sequence alignments
```{r eval = TRUE}
suppressPackageStartupMessages({
library(Biostrings)
  })
fasta_db <- readAAStringSet("oma_db")

@@

# WIF Family
## HMMER:hmmsearch
```{bash}
docker run --rm -ti -v $(pwd):/home/my_data  staphb/hmmer

hmmsearch --cut_ga --domtblout my_data/hmm_search/WIF_domtbl.out my_data/WIF.hmm my_data/oma_db > /dev/null
```

@@

WIF_hmmsearch <- data.table::fread(cmd = "grep -v '^#' hmm_search/WIF_domtbl.out | awk -v OFS='\t'  '{for(i=1;i<=23;i++) printf $i\"\t\"; print \"\"}'",  header = FALSE, sep = "\t",  fill=TRUE)

names(WIF_hmmsearch)[1:23] <- second_line_fields
```

## Multiple Sequence Alignment
```{r}
# filter for the hmmsearch env_coord sequences
fasta_db_fil_with_len <- fasta_db[names_fst_db %in% WIF_hmmsearch$target_name]

WIF_hmmsearch <- WIF_hmmsearch %>% 
  filter(!duplicated(target_name)) %>% 
  select(target_name, from_env_coord, to_env_coord) %>% 
  arrange(target_name)

identical(fasta_db_fil_with_len %>% names %>% str_remove(" .+"),
          WIF_hmmsearch$target_name)

#fasta_db_fil_with_len %>% head %>% Biostrings::as.list() %>% Biostrings::AAStringSet()
fasta_db_fil_with_len <- pmap(list(fasta_db_fil_with_len %>% as.list, 
          WIF_hmmsearch %>% select(target_name, from_env_coord) %>% deframe %>% as.list(),
          WIF_hmmsearch %>% select(target_name, to_env_coord) %>% deframe %>% as.list()),
     subseq) %>% Biostrings::AAStringSet()
```

## perform MSA with MUSCLE hmmsearch
```{r}
## env_coord
hmmsearch_env_coord_muscle <- msaMuscle(fasta_db_fil_with_len, type = "protein")
hmmsearch_env_coord_bio2mds <- msa::msaConvert(hmmsearch_env_coord_muscle, type = "bios2mds::align")
export.fasta(hmmsearch_env_coord_bio2mds, outfile = "WIF_hmmsearch_env_coord_msa_muscle.fa", ncol = 60, open = "w")
```
check the sign_seqs_msa_muscle.fa file for the alignments

@@

iqtree -s my_data/WIF_hmmsearch_env_coord_msa_muscle.fa -nt 4 -mset LG -pre my_data/my_hmmsearch_tree/WIF_hmmsearch_env_coord_tree

@@

>>>>

## import fasta file with all sequences
import the sequences and filter them for the 10 best hits from the previous results in order to make the multiple sequence alignments
```{r eval = TRUE}
suppressPackageStartupMessages({
library(Biostrings)
  })
fasta_db <- readAAStringSet("oma_db")
fasta_db <- fasta_db[sort(names(fasta_db))]

@@

# Searching for the Hs PORCN, protein-cysteine N-palmitoyltransferase porcupine isoform X5 [Homo sapiens]
## the Fasta file
```{bash}
cat porcn_hs.fasta 
>Hs|PORCN  gi|530421517|ref|XP_005272694.1| gi|530421517|ref|XP_005272694.1| PREDICTED: protein-cysteine N-palmitoyltransferase porcupine isoform X5 [Homo sapiens]
MYAIPSPSDWTVHSTGETEVLHSLQVTGCACACRTPTCTQLVAQSDVPDFSRLPPVLQYR
ASTWAGQPQTEISIHLAIHPWGSAMATFSRQEFFQQLLQGCLLPTAQQGLDQIWLLLAIC
LACRLLWRLGLPSYLKHASTVAGGFFSLYHFFQLHMVWVVLLSLLCYLVLFLCRHSSHRG
VFLSVTILIYLLMGEMHMVDTVTWHKMRGAQMIVAMKAVSLGFDLDRGEVGTVPSPVEFM
GYLYFVGTIVFGPWISFHSYLQAVQGRPLSCRWLQKVARSLALALLCLVLSTCVGPYLFP
YFIPLNGDRLLRNKKRKARGTMVRWLRAYESAVSFHFSNYFVGFLSEATATLAGAGFTEE
KDHLEWDLTVSKPLNVELPRSMVEVVTSWNLPMSYWLNNYVFKNALRLGTFSAVLVTYAA
SALLHGFSFHLAAVLLSLAFITYVEHVLRKRLARILSACVLSKRCPPDCSHQHRLGLGVR
ALNLLFGALAIFHLAYLGSLFDVDVDDTTEEQGYGMAYTVHKWSELSWASHWVTFGCWIF
YRLIG
```

## PFAM search
we search the sequence on the PFAM site and we get the MBOAT, 
membrane-bound O-acyltransferase family as result.
We download the model for the MBOAT

## HMMER:hmmsearch
```{bash}
docker run --rm -ti -v $(pwd):/home/my_data  staphb/hmmer

cd /home

hmmsearch --cut_ga --domtblout my_data/hmm_search/MBOAT_domtbl.out my_data/MBOAT.hmm my_data/oma_db > /dev/null
```

@@

MBOAT_hmmsearch <- data.table::fread(cmd = "grep -v '^#' hmm_search/MBOAT_domtbl.out | awk -v OFS='\t'  '{for(i=1;i<=23;i++) printf $i\"\t\"; print \"\"}'",  header = FALSE, sep = "\t",  fill=TRUE)

names(MBOAT_hmmsearch)[1:23] <- second_line_fields
```

## Multiple Sequence Alignment
```{r}
# filter for the hmmsearch env_coord sequences
fasta_db_fil_with_len <- fasta_db[names_fst_db %in% MBOAT_hmmsearch$target_name]

MBOAT_hmmsearch <- MBOAT_hmmsearch %>% 
  filter(!duplicated(target_name)) %>% 
  select(target_name, from_env_coord, to_env_coord) %>% 
  arrange(target_name)

identical(fasta_db_fil_with_len %>% names %>% str_remove(" .+"),
          MBOAT_hmmsearch$target_name)

#fasta_db_fil_with_len %>% head %>% Biostrings::as.list() %>% Biostrings::AAStringSet()
fasta_db_fil_with_len <- pmap(list(fasta_db_fil_with_len %>% as.list, 
          MBOAT_hmmsearch %>% select(target_name, from_env_coord) %>% deframe %>% as.list(),
          MBOAT_hmmsearch %>% select(target_name, to_env_coord) %>% deframe %>% as.list()),
     subseq) %>% Biostrings::AAStringSet()
```

## perform MSA with MUSCLE hmmsearch
```{r}
## env_coord
hmmsearch_env_coord_muscle <- msaMuscle(fasta_db_fil_with_len, type = "protein")
hmmsearch_env_coord_bio2mds <- msa::msaConvert(hmmsearch_env_coord_muscle, type = "bios2mds::align")
export.fasta(hmmsearch_env_coord_bio2mds, outfile = "MBOAT_hmmsearch_env_coord_msa_muscle.fa", ncol = 60, open = "w")
```
check the sign_seqs_msa_muscle.fa file for the alignments

@@

iqtree -s my_data/MBOAT_hmmsearch_env_coord_msa_muscle.fa -nt 4 -mset LG -pre my_data/my_hmmsearch_tree/MBOAT_hmmsearch_env_coord_tree

@@

# hmmmbuild
```{bash}
hmmbuild --amino --cpu 4  my_data/NOTUM_try my_data/hmmsearch_env_coord_msa_muscle.fa 
hmmbuild --amino --cpu 4  my_data/WIF_try my_data/WIF_hmmsearch_env_coord_msa_muscle.fa 
hmmbuild --amino --cpu 4  my_data/wnt_try my_data/wnt_hmmsearch_env_coord_msa_muscle.fa
